Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-05-01 Chapter-1/1-Theory.tex
s

Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories as a function of increasing neutrino energy: quasi elastic (QE), resonant (RES), and deep inelastic (DIS) scattering. All current and forthcoming oscillation experiments live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into RES and the energy  where RES  transitions into DIS. 
 Schematically, neutrino and antineutrino QE charge current scattering refers to the process $\nu_l n \rightarrow l^- p$ and $\bar\nu_l p \rightarrow l^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction, leaving the target nucleus in its ground state. %of a neutrino (or antineutrino) with a nucleon in the target material.  
 Resonant scattering refers to an inelastic collision producing a nucleon excited state ($\Delta$,N*) -- the resonance, indeed -- which then quickly decays, most often to a nucleon and single-pion final state. DIS refers to the head-on collision between the neutrino and a parton inside the nucleon, producing hadronization and subsequent  abundant production of mesons and nucleons.  In addition to such interactions between the neutrino and a single component of the nucleus, neutrinos can also interact with the nucleus as a whole, albeit more rarely, a well documented process called coherent meson production scattering \cite{Vilain1993}; the signature of such process is the production of a distinctly forward-scattered single meson final state, most often a pion.  This simple picture of neutrino interactions works rather well for scattering 
off of light nuclear targets, such as the H$_2$ and D$_2$ of bubble chamber experiments \cite{RevModPhys.84.1307}, but the complexity of the nuclear structure for heavier nuclei such as argon complicates this model.

As we will discuss in Chapter \ref{ch:2}, the properties of argon make it a good candidate for interacting medium in neutrino experiments; in particular the density of its interaction centers  augments the yield of neutrino interactions and allows for relatively compact detectors. Though, the choice of a relatively heavy nuclear target comes at the cost of enhancing  nuclear effects which modify the kinematic and final state of the neutrino interaction products.

Nuclear effects can potentially affect the neutrino event rates, nucleon emission, neutrino energy reconstruction, and the neutrino/antineutrino ratios, carrying deep implications for oscillation experiments. 
Even in the case of ``simple" QE scattering, intra-nuclear hadron rescattering and correlation effects between the target nucleons can cause the ejection of additional nucleons in the final state, modifying the final state kinematics and topology. In case of resonant and DIS scattering, the hadronic interactions of meson and nucleons produced in the decay of the resonance or during hadronization  complicate this picture even more.
A large source of uncertainty in modeling  nuclear effects in neutrino interactions come from mesons interactions (and re-interactions) in the nucleus, e.g., pion re-scattering, charge exchange, and absorption.

@@

\subsection{Pion-Argon Total Hadronic Cross Section}
This section outlines the importance of the pion-argon total hadronic cross section in the context of the current and upcoming liquid argon neutrino experiments, SBN and DUNE. %We start by discussing the measurement in the context of neutrino interaction searches. % and of light mesons interaction with nuclei studies. 
 We describe the signal signature and historic measurements of pion-nucleus cross section, as well as the implementation of this cross sections in the current version of the simulation package used by LArIAT.

\subsubsection{$\pi^{-}$Ar Cross Section in the Context of Neutrino Searches}
As outlined in \ref{ch:NuInt}, neutrino experiments use the products of neutrino interactions to identify the energy and flavor of the incoming neutrino.
Pions are a common product of neutrino interaction, especially in resonant scattering, DIS and coherent pion production. For neutrino experiments in argon, there are two main reasons why understanding pion hadronic interactions with argon is important: to model the behavior of the pion inside the nucleus struck by the neutrino and to model the behavior of the pion during its propagation inside the detector medium.

Assumptions on the nuclear models and on the interaction of hadrons inside the nucleus performed at the level of the neutrino event generator bridge the measurement of the products of a neutrino interaction to the reconstruction of the neutrino energy and flavor. Thus, understanding pion hadronic interactions with the nucleus is particularly important to model correctly resonant, DIS and coherent pion production in neutrino interactions, where the presence of pions in the nucleus is abundant. For example, in case of resonant scattering,
\begin{equation}
\nu_l + N \rightarrow l + \Delta/N^* \rightarrow l+ \pi + N', 
\end{equation}
the $\Delta$ and $N^*$ and excited states will decay hadronically in matters of $\sim10^{-24}$~s inside the nucleus producing pions which will bounce within the nuclear medium. The decay modes for the lower mass $\Delta$ (1232) and $N^*$(1440)  are listed in table \ref{tab:decaysdelta}.
\begin{table}[b]
\centering
\begin{tabular}{|l|l|l|}
\hline
Resonance  & Decay Mode & Lifetime (s)\\
\hline
    &  &  \\
$\Delta$ (1232) 3/2$^+$  & \begin{tabular}{@{}l@{}} $\Delta^{++} \text{(uuu)} \rightarrow p \pi^+$ \\$\Delta^{+} \text{(uud)} \rightarrow n \pi^+$  \\ $\Delta^{+} \text{(uud)} \rightarrow p \pi^0$\\$\Delta^{0} \text{(udd)} \rightarrow n \pi^0$ \\ $\Delta^{0} \text{(udd)} \rightarrow p \pi^-$ \\$\Delta^{-} \text{(ddd)} \rightarrow n \pi^-$  \end{tabular}
 & $\sim5.6 \times10^{-24}$\\
     &  &  \\
\hline
    &  &  \\
$N^*$ (1440) 1/2$^+$     & \begin{tabular}{@{}l@{}}$N^* \rightarrow N \pi$ \\$N^* \rightarrow N \pi\pi$\end{tabular} &   $\sim2.2 \times10^{-24}$\\
    &  &  \\
\hline
\end{tabular}
\caption{Main decay modes of the lightest Delta resonance and Nucleon excited state.}
\label{tab:decaysdelta}
\end{table}

@@

The key elements of a neutrino event generators for resonance and DIS events are the nuclear model and the hadron treatment (both production and transportation).  We illustrate here the conceptual basis of the GENIE Neutrino Generator~\cite{Andreopoulos:2009rq} as an example, since GENIE is one the most popular event generators for liquid argon experiments.
For example, the nuclear model used by GENIE for all processes is a relativistic Fermi gas (RFG) modified to incorporate short range nucleon-nucleon correlations \cite{PhysRevD.24.1400}. This means that the initial momentum and binding energy of the struck nucleon is determined by assuming nucleons inside the nucleus are quasi-free, acting independently in the mean field of the nucleus. For A $ >$ 20 like argon, the 2-parameter Woods-Saxon shell model for density function is used.  The GENIE module INTRANUKE \cite{Juszczak2006} is used to simulate the final-state interactions (FSI) which is the hadron re-interaction inside the nucleus. This module places the outgoing particles in the nucleus and propagates them using the ``hA model".  In the INRANUKE hA model, hadrons can undergo at most one FSI per event. When possible, external hadron-nucleus scattering data are used to tune INTRANUKE. Since no data is available for Argon, GENIE uses an interpolation of data from heavier and lighter nuclei for the pion-argon cross section leading to big uncertainties in the INTRANUKE module.

Once the pion has left the target nucleus, the pion-argon hadronic cross section plays an important role in the pion transportation inside the argon medium: processes like pion absorption with emission of nucleons or pion charge exchange can greatly modify the topology of a neutrino interactions in the detector and lead to errors in the event classification.
Being able to reconstruct the details of pions inside the detector is an imperative for modern liquid argon neutrino experiment to achieve the design resolution for their key physics measurements.

\subsubsection{$\pi^{-}$Ar Hadronic Interaction: Signal Signatures}
Strong hadronic interaction models \cite{9780198520085,9780471779957} predict the pion interaction processes with argon in the [100 - 1200] MeV energy range. 
The total hadronic $\pi^{-}$-Ar interaction cross section is defined as the one related to the single process driven only by the strong interaction which is dominant in the considered energy range. 
In measuring the ``total" cross section, we include both the elastic and  reaction channels, regardless of the final state, 
\begin{equation}
\sigma_{Tot} = \sigma_{Elastic}+ \sigma_{Reaction};
\end{equation}
the reaction channel is further characterizes by several exclusive channels with defined topologies,
\begin{equation}
\sigma_{Reaction} = \sigma_{Inelastic} + \sigma_{abs} + \sigma_{chex}+ \sigma_{\pi prod}.
\end{equation}

A summary of the pion final states in order of pion multiplicity for the reaction channel is given in table \ref{tab:pionFinal}. Pion capture and pion decay at rest dominate the cross section under 100 MeV. We define pion capture as the process determining the formation of a pionic atom and the subsequent pion's end of life. Stopping negative pions can form pionic argon, where the negative pion plays the role of an orbital electron. Since the pion mass is two orders of magnitude greater than the electron mass, the spatial wave form of the pion will overlap more with the nucleus compared to the electron case. After the electromagnetic formation of the pionic atom, the pion will be quickly absorbed by the nucleus, which is put in an excited state. The nucleus then de-excites with the emission of low energy nucleons and photons.  Pion capture is predominant compared to pion decay, the other important process for very low energy pions. The decay of a pion is governed by the weak force; the pion decay life time is $\tau_\pi = 2.6 \times 10^{-8}$ s and the main decay mode is  $\pi^-\rightarrow \mu^- + \bar\nu_\mu$ (BR 99.98\%). %; for pion decay at rest, the muon is emitted with an energy of $\sim5$ MeV and it decays into $\mu^-\rightarrow e^- + \bar\nu_e + \nu_\mu$ in $\tau_\mu = 2.3 $ $\mu$s. 
 Since pion capture can be considered an electromagnetic process and pion decay is a weak process, this energy region is purposely excluded from the hadronic cross section measurement.

@@

\subsubsection{Previous measurements: Lighter and Heavier Nuclei}
Many experiments with pion beams have studied the hadronic interaction of pions on light and heavy materials, such as  He, Li, C, Fe, Pb \cite{PhysRevC.14.635}. However, data on argon are rare: the total differential hadronic cross section has never been measured before on argon. Simulation packages like Geant4 base their pion transportation for argon on data from lighter and heavier nuclei: the goal of LArIAT's dedicated measurement on argon is to bridge this gap in data, thus reducing the uncertainties related to pions interaction in argon in both neutrino event generators and in simulation packages of pion transportation.

@@

\subsubsection{Negative Pion Interaction Cross Section in Simulation Packages}
LArIAT uses Geant4 as the default simulation package. In particular, pions (and kaons) transportation is achieved through the Geant4 FTFP\_BERT physics list. In this physics list, Geant4 uses the Bertini cascade model  \cite{Wright:2015xia}  to simulate the products of the pion-nucleus interaction as well as the secondary hadrons re-interactions inside the target nucleus (intra-nuclear cascade).  The target nucleus is represented as a continuous gas where the nuclear potential follows concentrical shells whose depths approximate the Woods-Saxon shape.  The CERN-HERA compilations \cite{Compilation1,Compilation2} of hadron-nucleon interaction data is the data base used for the decision making process after the cascade is invoked. The cross section model determines if the pion interacts, the eventual type of interaction and the interaction multiplicity. For  hadron projectiles with energy less than 20 GeV, Geant4 reports  the uncertainty on the cross section model to be about the size of the error bars on the data used, or about 10\%, increasing to 20-30\% in energy regions where data is sparse.

The relevance of the GENIE generator for neutrino physics and its basic working principles have been outlined earlier in this section. Given GENIE's modularity, information on hadron-nucleus interactions can be extracted from the  INTRANUKE module and directly compared against the Geant4 predictions.
The work in \cite{Nutini2015}  reviews  the current status of negative and positive pion simulation in Geant4 and GENIE for $^{12}$C, $^{56}$Fe, and $^{40}$Ca. From that work, we report the results for $^{12}$C in Figure \ref{fig:pionC} as it allows a direct comparison between Geant4, GENIE and data.  Geant4 predictions for $\pi^-$  on Carbon are in good agreement with data over all the spectrum, while GENIE predictions seem to show some features at around 500 MeV and 900 MeV, maybe due to higher resonances in the hA model. From the same work, we also report the negative pion cross section on $^{40}$Ca in Figure \ref{fig:pionCa}, since this is the nuclear medium closest argon with some available data. The predictions from both Geant4 and GENIE agree with data in the high energy region; the Geant4 and GENIE predictions diverge in the resonance region, where data is not available. These few examples highlight how cross section data for the specific nucleus considered in the neutrino experiments is fundamental to inform the Monte Carlo simulation.

>>>>

Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories into three categories whose contributions change as a function of increasing neutrino energy:: quasi elastic (QE), resonant (RES), and deep inelastic (DIS) scattering. All current and forthcoming oscillation experiments live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into RES and then into DIS. 
For scattering off free nucleons, neutrino and antineutrino QE charge current scattering refers to the process $\nu_l n \rightarrow l^- p$ and $\bar\nu_l p \rightarrow l^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction. %of a neutrino (or antineutrino) with a nucleon in the target material.  
 Resonant scattering refers to an inelastic collision producing a nucleon excited state ($\Delta$,N*) -- the resonance -- which then quickly decays, most often to a nucleon and single-pion final state. DIS refers to the head-on collision between the neutrino and a parton inside the nucleon, producing hadronization and subsequent  abundant production of mesons and nucleons.  In addition to such interactions between the neutrino and a single component of the nucleus, neutrinos can also interact with the nucleus as a whole, albeit more rarely, a well documented process called coherent meson production scattering \cite{Vilain1993}; the signature of such process is the production of a distinctly forward-scattered single meson final state, most often a pion.  This simple picture of neutrino interactions works rather well for scattering 
off of light nuclear targets, such as the H$_2$ and D$_2$ of bubble chamber experiments \cite{RevModPhys.84.1307}, but the complexity of the nuclear structure for heavier nuclei such as argon complicates this model.

As we will discuss in Chapter \ref{ch:2}, the properties of argon make it a good candidate for an interacting medium in neutrino experiments; in particular the density of its interaction centers  increases the yield of neutrino interactions and allows for relatively compact detectors. Though, the choice of a relatively heavy nuclear target comes at the cost of enhancing  nuclear effects which modify the kinematic and final state of the neutrino interaction products.

Nuclear effects can potentially affect neutrino event rates, final state particle emission, neutrino energy reconstruction, and the neutrino/antineutrino ratios, carrying deep implications for oscillation experiments. 
Even in the case of ``simple" QE scattering, intra-nuclear hadron rescattering and correlation effects between the target nucleons can cause the ejection of additional nucleons in the final state, modifying the final state kinematics and topology. In the case of resonant and DIS scattering, the hadronic interactions of meson and nucleons produced in the decay of the resonance or during hadronization  complicate this picture even more.
A large source of uncertainty in modeling  nuclear effects in neutrino interactions come from mesons interactions (and re-interactions) in the nucleus, e.g., pion re-scattering, charge exchange, and absorption.

@@

\subsection{Pion-Argon Total Hadronic Cross Section}
This section outlines the importance of the pion-argon total hadronic cross section in the context of the current and upcoming liquid argon neutrino experiments, SBN and DUNE. %We start by discussing the measurement in the context of neutrino interaction searches. % and of light mesons interaction with nuclei studies. 
 We describe the signal signature and historic measurements of pion-nucleus cross section, as well as the implementation of these cross sections in the current version of the simulation package used by LArIAT.

\subsubsection{$\pi^{-}$Ar Cross Section in the Context of Neutrino Searches}
As outlined in \ref{ch:NuInt}, neutrino experiments use the products of neutrino interactions to identify the energy and flavor of the incoming neutrino.
Pions are a common product of neutrino interaction, especially in resonant scattering, DIS and coherent pion production. For neutrino experiments in argon, there are two main reasons why understanding pion hadronic interactions with argon is important: to model the behavior of the pion inside the target nucleus  and to model the behavior of the pion during its propagation inside the detector medium.

Assumptions on the nuclear modeling and on the interaction of hadrons inside the nucleus performed at the level of the neutrino event generator bridge the measurement of the products of a neutrino interaction to the reconstruction of the neutrino energy and flavor. Thus, understanding pion hadronic interactions with the nucleus is particularly important to model correctly resonant, DIS and coherent pion production in neutrino interactions. For example, in case of resonant scattering,
\begin{equation}
\nu_l + N \rightarrow l + \Delta/N^* \rightarrow l+ \pi + N', 
\end{equation}
the $\Delta$ and $N^*$ and excited states will decay hadronically in matters of $\sim10^{-24}$~s inside the nucleus producing pions which which will have many chances to re-interact as they exit the target medium. The decay modes for the lower mass $\Delta$ (1232) and $N^*$(1440)  are listed in table \ref{tab:decaysdelta}.
\begin{table}[b]
\centering
\begin{tabular}{|l|l|l|}
\hline
Resonance  & Decay Mode & Lifetime (s)\\
\hline
    &  &  \\
$\Delta$ (1232) 3/2$^+$  & \begin{tabular}{@{}l@{}} $\Delta^{++} \text{(uuu)} \rightarrow p \pi^+$ \\$\Delta^{+} \text{(uud)} \rightarrow n \pi^+$  \\ $\Delta^{+} \text{(uud)} \rightarrow p \pi^0$\\$\Delta^{0} \text{(udd)} \rightarrow n \pi^0$ \\ $\Delta^{0} \text{(udd)} \rightarrow p \pi^-$ \\$\Delta^{-} \text{(ddd)} \rightarrow n \pi^-$  \end{tabular}
 & $\sim5.6 \times10^{-24}$\\
     &  &  \\
\hline
    &  &  \\
$N^*$ (1440) 1/2$^+$     & \begin{tabular}{@{}l@{}}$N^* \rightarrow N \pi$ \\$N^* \rightarrow N \pi\pi$\end{tabular} &   $\sim2.2 \times10^{-24}$\\
    &  &  \\
\hline
\end{tabular}
\caption{Main decay modes of the lightest Delta resonance and Nucleon excited state.}
\label{tab:decaysdelta}
\end{table}

@@

The key elements of a neutrino event generators for resonance and DIS events are the nuclear model and the hadron treatment (both production and transportation).  We illustrate here the conceptual basis of the GENIE Neutrino Generator~\cite{Andreopoulos:2009rq} as an example, since GENIE is one the most popular event generators for liquid argon experiments.
For example, the nuclear model used by GENIE for all processes is a Relativistic Fermi Gas (RFG) model modified to incorporate nucleon-nucleon correlations \cite{PhysRevD.24.1400}. This means that the initial momentum and binding energy of the struck nucleon is determined by assuming nucleons inside the nucleus are quasi-free, acting independently in the mean field of the nucleus. For A $ >$ 20 such as argon, the 2-parameter Woods-Saxon shell model for density function is used.  The GENIE module INTRANUKE \cite{Juszczak2006} is used to simulate final-state interactions (FSI) which model  hadron re-interactions inside the nucleus. This module places the outgoing particles in the nucleus and propagates them using the ``hA model".  In the INRANUKE hA model, hadrons can undergo at most one FSI per event. When possible, external hadron-nucleus scattering data are used to tune INTRANUKE. Since no data is available for Argon, GENIE uses an interpolation of data from heavier and lighter nuclei for the pion-argon cross section leading to large (10?s of \%) resultant uncertainties in the INTRANUKE module.

Once the pion has left the target nucleus, the pion-argon hadronic cross section also plays an important role in the pion transportation inside the argon medium: processes such as pion absorption or pion charge exchange can greatly modify the topology of a neutrino interactions in the detector and lead to significant modifications in the event classification.
Being able to reconstruct the details of pions inside the detector is an imperative for modern liquid argon neutrino experiment to achieve the design resolution for their key physics measurements.
%For example, if a pi+ charge exchanges into a pi0, an event which would not have been a background to a nue search now becomes one. So you need to know with what frequency this happens in a given nucleus for a given energy pion. Also, if a pion gets absorbed in the target nucleus, this energy is  transferred into nucleons, some of which can be  invisible (e.g., neutrons) to the  detector thus leading to a underestimate of  incoming neutrino energy.

\subsubsection{$\pi^{-}$Ar Hadronic Interaction: Signal Signatures}
Strong hadronic interaction models \cite{9780198520085,9780471779957} predict the pion interaction processes with argon in the [100 - 1200] MeV energy range. 
The total hadronic $\pi^{-}$-Ar interaction cross section is defined as the one related to the single process driven only by the strong interaction which is dominant in the considered energy range. 
In measuring the ``total" cross section, we include both the elastic and  reaction channels, regardless of the final state, 
\begin{equation}
\sigma_{Tot} = \sigma_{Elastic}+ \sigma_{Reaction};
\end{equation}
the reaction channel is further characterized by several exclusive channels with defined topologies,
\begin{equation}
\sigma_{Reaction} = \sigma_{Inelastic} + \sigma_{abs} + \sigma_{chex}+ \sigma_{\pi prod}.
\end{equation}

A summary of the pion final states in order of pion multiplicity for the reaction channel is given in table \ref{tab:pionFinal}. Pion capture and pion decay at rest dominate the cross section under 100 MeV. We define pion capture as the process determining the formation of a pionic atom and the subsequent pion's end of life. Stopping negative pions can form pionic argon, where the negative pion plays the role of an orbital electron. Since the pion mass is two orders of magnitude greater than the electron mass, the spatial wave form of the pion will overlap more with the nucleus compared to the electron case. After the electromagnetic formation of the pionic atom, the pion will get quickly absorbed by the nucleus, which is put in an excited state. The nucleus then de-excites with the emission of low energy nucleons and photons.  Pion capture is dominant compared to pion decay, the other important process for very low energy pions. The decay of a pion is governed by the weak force; the pion decay life time is $\tau_\pi = 2.6 \times 10^{-8}$ s and the main decay mode is  $\pi^-\rightarrow \mu^- + \bar\nu_\mu$ (BR 99.98\%). %; for pion decay at rest, the muon is emitted with an energy of $\sim5$ MeV and it decays into $\mu^-\rightarrow e^- + \bar\nu_e + \nu_\mu$ in $\tau_\mu = 2.3 $ $\mu$s. 
 Since pion capture can be considered an electromagnetic process and pion decay is a weak process, this energy region is purposely excluded from the hadronic cross section measurement.

@@

\subsubsection{Previous measurements: Lighter and Heavier Nuclei}
Many experiments with pion beams have studied the hadronic interaction of pions on light and heavy materials, such as  He, Li, C, Fe, Pb \cite{PhysRevC.14.635}. However, data on argon are rare: the total differential hadronic cross section has never been measured before on argon. Simulation packages such as Geant4 base their pion transportation for argon on data from lighter and heavier nuclei: the goal of LArIAT's dedicated measurement on argon is to bridge this gap in data, thus reducing the uncertainties related to pion interactions in argon in both neutrino event generators and in simulation packages of pion transportation.

@@

\subsubsection{Negative Pion Interaction Cross Section in Simulation Packages}
LArIAT uses Geant4 as the default simulation package. In particular, pions (and kaons) transportation is achieved through the Geant4 FTFP\_BERT physics list. In this physics list, Geant4 uses the Bertini cascade model  \cite{Wright:2015xia}  to simulate the products of the pion-nucleus interaction as well as secondary hadronic re-interactions inside the target nucleus (intra-nuclear cascade).  The target nucleus is represented as a continuous gas where the nuclear potential follows concentrical shells whose depths approximate the Woods-Saxon shape.  The CERN-HERA compilations \cite{Compilation1,Compilation2} of hadron-nucleon interaction data is the data base used for the decision making process after the cascade is invoked. The cross section model determines if the pion interacts, the eventual type of interaction and the interaction multiplicity. For  hadron projectiles with energy less than 20 GeV, Geant4 reports  the uncertainty on the cross section model to be about the size of the error bars on the data used, or about 10\%, increasing to 20-30\% in energy regions where data is sparse.

The relevance of the GENIE generator for neutrino physics and its basic working principles have been outlined earlier in this section. Given GENIE's modularity, information on hadron-nucleus interactions can be extracted from the  INTRANUKE module and directly compared against the Geant4 predictions.
The work in \cite{Nutini2015}  reviews  the current status of negative and positive pion simulation in Geant4 and GENIE for $^{12}$C, $^{56}$Fe, and $^{40}$Ca. From that work, we report the results for $^{12}$C in Figure \ref{fig:pionC} as it allows a direct comparison between Geant4, GENIE and and pion re-scattering data.  Geant4 predictions for $\pi^-$  on Carbon are in good agreement with data over  the entire spectrum spectrum, while GENIE predictions seem to show some features at around 500 MeV and 900 MeV, maybe due to higher resonances in the hA model. From the same work, we also report the negative pion cross section on $^{40}$Ca in Figure \ref{fig:pionCa}, since this is the nuclear medium closest to argon. The predictions from both Geant4 and GENIE agree with data in the high energy region; the Geant4 and GENIE predictions diverge in the resonance region, where data is not available. These few examples highlight how cross section data for the specific nucleus considered in the neutrino experiments is fundamental to inform the Monte Carlo simulation.
